Normalization by Evaluation for Call-by-Push-Value and Polarized Lambda-Calculus

### 2019-02-13 fscd19/fscd19.tex
[ fscd19 ] removed some fboxes

The category $\Cxt$ allows us to consider $A{\in}\_$ as a presheaf over
$\Cxt\op$ for any $A$, witnessed by \fbox{$\treindex : \Gamma \sublist
  \Delta \to A \in \Gamma \to A \in \Delta$}, which is the morphism
part of functor $A{\in}\_$ from $\Cxt$ to $\Set$, mapping object $\Gamma$ to the
set $A \in \Gamma$ of the indices of $A$ in $\Gamma$.
The associated naturality laws \fbox{$\treindex\,\tid\,x \equiv x$}
and \fbox{$\treindex\,\tau_2\,(\treindex\,\tau_1\,x) \equiv
\treindex\,(\tau_1 \rcomp \tau_2)\,x$} hold propositionally.

@@

% that does not guarantee any relation between its input term $t$ and
% output normal form $u$.
%
% \emph{Normalization by evaluation} (NbE) decomposes normalization as
% \fbox{\(
%   \norm[\Gamma] A (t : A \from \Gamma)
%   = \down[\Gamma] A \dent t {\idenv\Gamma}
% \)}
% which is
% \emph{evaluation}
% \nofbox{$\dent \_ \_ : (t : A \from \Gamma) \to \Den \Gamma \Delta \to
%   \Den A \Delta$}
% in the identity environment \nofbox{$\idenv\Gamma : \Den \Gamma \Gamma$}
% followed by
% \emph{reification}
% \nofbox{$\downG A : \Den A \Gamma \to \Nf A \Gamma$}.
\emph{Normalization by evaluation} (NbE) decomposes normalization as
\fbox{\(
%  \norm[\Gamma] A (t : A \from \Gamma)
  \tnorm\, (t : A \from \Gamma)
  = \down A \dent t {\idenv\Gamma}
\)}
which is
\emph{evaluation}
\nofbox{$\dent \_ \_ : (t : A \from \Gamma) \to \den \Gamma \todot
  \den A$}
in the identity environment \nofbox{$\idenv\Gamma : \den \Gamma \Gamma$}
followed by
\emph{reification}
\nofbox{$\down A : \den A \todot \tNf A$}.
The role of evaluation is to produce from a term the corresponding
semantic (\ie, meta-theoretic) function, which is finally reified to a normal form.
Since we are evaluating open terms $t$, we need to supply an environment
$\idenv\Gamma$ which will map the free indices of $t$ to corresponding
\emph{unknowns}.  To accommodate unknowns in the semantics, types $A$
are mapped to presheaves $\den A$ (rather than just sets), and in
particular each base type $\tybase$ is mapped to the presheaf
$\tNe\,\tybase$ with the intention that the neutrals take the role of
the unknowns.  The mapping $\up A : \tNe\,A \todot \den A$ from
neutrals to unknowns is called \emph{reflection}, and defined mutually
with reification by induction on type $A$.

>>>>

The category $\Cxt$ allows us to consider $A{\in}\_$ as a presheaf over
$\Cxt\op$ for any $A$, witnessed by $\treindex : \Gamma \sublist
  \Delta \to A \in \Gamma \to A \in \Delta$, which is the morphism
part of functor $A{\in}\_$ from $\Cxt$ to $\Set$, mapping object $\Gamma$ to the
set $A \in \Gamma$ of the indices of $A$ in $\Gamma$.
The associated naturality laws $\treindex\,\tid\,x \equiv x$
and $\treindex\,\tau_2\,(\treindex\,\tau_1\,x) \equiv
\treindex\,(\tau_1 \rcomp \tau_2)\,x$ hold propositionally.

@@

% that does not guarantee any relation between its input term $t$ and
% output normal form $u$.
%
% \emph{Normalization by evaluation} (NbE) decomposes normalization as
% \fbox{\(
%   \norm[\Gamma] A (t : A \from \Gamma)
%   = \down[\Gamma] A \dent t {\idenv\Gamma}
% \)}
% which is
% \emph{evaluation}
% \nofbox{$\dent \_ \_ : (t : A \from \Gamma) \to \Den \Gamma \Delta \to
%   \Den A \Delta$}
% in the identity environment \nofbox{$\idenv\Gamma : \Den \Gamma \Gamma$}
% followed by
% \emph{reification}
% \nofbox{$\downG A : \Den A \Gamma \to \Nf A \Gamma$}.
\emph{Normalization by evaluation} (NbE)
\fbox{\(
%  \norm[\Gamma] A (t : A \from \Gamma)
  \tnorm\, (t : A \from \Gamma)
  = \down A \dent t {\idenv\Gamma}
\)}
decomposes normalization into
\emph{evaluation}
\nofbox{$\denty\_ : (t : A \from \Gamma) \to \den \Gamma \todot
  \den A$}
in the identity environment \nofbox{$\idenv\Gamma : \den \Gamma \Gamma$}
followed by
\emph{reification}
\nofbox{$\down A : \den A \todot \tNf A$}.
The role of evaluation is to produce from a term the corresponding
semantic (\ie, meta-theoretic) function, which is finally reified to a normal form.
Since we are evaluating open terms $t$, we need to supply an environment
$\idenv\Gamma$ which will map the free indices of $t$ to corresponding
\emph{unknowns}.  To accommodate unknowns in the semantics, types $A$
are mapped to presheaves $\den A$ (rather than just sets), and in
particular each base type $\tybase$ is mapped to the presheaf
$\tNe\,\tybase$ with the intention that the neutrals take the role of
the unknowns.  The mapping $\up A : \tNe\,A \todot \den A$ from
neutrals to unknowns is called \emph{reflection}, and defined mutually
with reification by induction on type $A$.
